---
title: "readme"
author: "Edward Lee"
date: "`r Sys.Date()`"
output: md_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction 

This is the repository for the paper title "Thawed Plasma vs Liquid Plasma: Tracking Coagulation Factor Activity Changes During Storage" by Yurtsever et al.
We have included the coagulation factor activity data (fibrinogen, factor V, factor VII, factor VIII, protein C, and protein S) measured in liquid plasma (on day 15, 26, and 27 of storage) and in thawed plasma (on day 5).

## Data
The data set is included in the `figures` folder.
It is provided in both .xlsx and .rds formats for Excel and R users respectively.
The data has the following columns:

- `sample_id`: IDs of the plasma units used;
- `product`: describes if the plasma unit is either "liquid plasma" or "thawed plasma"
- `day`: the day of storage when the factor activity measurement was performed
- `factor`: which factor was measured
- `value`: the activity measurement; fibrinogen has the units mg/dL, and the others have the units %
- `comments`: any relevant comments regarding the measurements

## Code
The statistical models are coded in R and uses Stan (see [here](https://mc-stan.org)) for sampling the posteriors of the Bayesian models.
We use both Richard Mcelreath's rethinking package and directly coded Stan models in this study.
The R scripts are contained in the `scripts` folder, and the Stan code are contained in the `stan` folder.
The following scripts implement the various parts of the analysis:

- `lqp_coag_factors.R` runs the Student-t analysis to estimate the posterior distribution of coagulation factor activity for each product at each time point
- `compare_lqp_tp5.R` runs the Bayesian two-sample t test analysis
- `regression_all.R` runs the multilevel piecewise linear regression analysis
- `student-t-analysis.R` contains the analysis of comparing a fixed nu vs a free nu when fitting the Student-t likelihoods

Running the code requires installation of `tidyverse`, `rethinking`, `cmdstanr`, and `readxl` packages, which can be installed by running the following commands 
```{r eval = FALSE}
install.packages(c("tidyverse", "cmdstanr", "readxl"))
devtools::install_github("rmcelreath/rethinking")
```

